Ethereum: A Secure Decentralised Generalised Transaction Ledger

### 2014-04-17 Paper.tex
Conventions.

\begin{eqnarray}
\mathbb{S}_{t+1} & \equiv & \Pi(\mathbb{S}_t, B) \\
B & \equiv & (..., [ T_0, T_1, ... ]) \\
\Pi(\mathbb{S}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\mathbb{S}, T_0), T_1) ...)
\end{eqnarray}

@@

@@

\begin{description}
\item[nonce] A scalar value equal to the number of transactions sent from this address, or, in the case of contract accounts, the number of contract-creations made by this account.
\item[balance] A scalar value equal to the number of wei owned by this address.
\item[stateRoot] A 256-bit hash equal to the root node of a further trie structure that encodes the storage contents of the contract. Though a separate data structure, this trie is still stored using the same underlying state database. This trie takes the form as a simple mapping between domains of 256-bit values.
\item[codeHash] The hash of the EVM code of this contract---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction. All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
\end{description}

@@

the validity is determined as:
\begin{eqnarray}
S(T) & \neq & \varnothing \\ \wedge \quad \mathbb{S}[S(T)] & \neq & \varnothing \\ \wedge \quad T_n & = & \mathbb{S}[S(T)]_{nonce} \\ \wedge \quad g_0 & \leq & T_g \\ \wedge \quad v_0 & \leq & \mathbb{S}[S(T)]_{balance}
\end{eqnarray}

@@

We define the rollback state $\mathbb{S}_0$:
\begin{eqnarray}
\mathbb{S}_0 & \equiv & \mathbb{S} \quad \text{except:} \\
\mathbb{S}_0[S(T)]_{balance} & \equiv & \mathbb{S}[S(T)]_{balance} - v_0 \\
\mathbb{S}_0[S(T)]_{nonce} & \equiv & \mathbb{S}[S(T)]_{nonce} + 1
\end{eqnarray}

@@

After the message call or contract creation is processed, the state is finalised by refunding $g'$, the remaining gas, to the sender at the original rate. The Ether for the gas that was actually used is given to the miner, whose address is specified as the coinbase of the present block $B$. So we define the final state $\mathbb{S}'$ in terms of the provisional state $\mathbb{S}_P$:
\begin{eqnarray}
\mathbb{S}' & \equiv & \mathbb{S}_P \quad \text{except} \\
\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_P[s]_{balance} + g' T_p \\
\mathbb{S}'[m]_{balance} & \equiv & \mathbb{S}_P[m]_{balance} + (T_g - g' - g_d) T_p \\
m & \equiv & {B_H}_b
\end{eqnarray}

%In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
%\begin{eqnarray}
%\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_P[s]_{balance} + g
%\end{eqnarray}

@@

The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the SHA3 256-bit hash of the code, thus the mutated state becomes $\mathbb{S}^*$:
\begin{eqnarray}
\mathbb{S}^* & \equiv & \mathbb{S} \quad \text{except:} \\
\mathbb{S}^*[A(s, \mathbb{S}[s]_{nonce})] & \equiv & ( 0, v, \varnothing, \mathtt{\tiny SHA3}(\mathbf{b}) )
\end{eqnarray}

@@

We define $\mathbb{S}_1$, the rollback state as the original state but with the value transferred to the recipient:
\begin{eqnarray}
\mathbb{S}_1 & \equiv & \mathbb{S} \quad \text{except:} \\
\mathbb{S}_1[r]_{balance} & \equiv & \mathbb{S}_1[r]_{balance} + v
\end{eqnarray}

If the recipient account contains no code to be executed (i.e. it's not a contract), then we define the final state as being equivalent to the rollback state. In the case that the recipient contains code to be executed (i.e. that the account is a contract), then the contract's body code (identified as the fragment whose SHA3 hash is $\mathbb{S}[r]_{code}]$) is executed according to the execution model (see section \ref{ch:model}). Just as with contract creation, if the execution halts due to an exhausted gas supply, then no gas is refunded to the caller and the state is reverted to the point immediately prior to code execution.
\begin{eqnarray}
\mathbb{S}_P & \equiv & \begin{cases}
\mathbb{S}_1 & \text{if} \quad \mathbb{S}_1[r]_{code} = \varnothing \\
\mathbb{S}_1 & \text{if} \quad \mathbb{S}^{**} = \varnothing \\
\mathbb{S}^{**} & \text{otherwise}
\end{cases} \\
(\mathbb{S}^{**}, g', \mathbf{o}) & \equiv & \Xi(\mathbb{S}_1, g, I) \\
I_a & \equiv & a \\
I_o & \equiv & o \\
I_p & \equiv & p \\
I_\mathbf{d} & \equiv & \mathbf{d} \\
I_s & \equiv & s \\
I_v & \equiv & v \\
\text{Let} \quad \mathtt{\tiny SHA3}(I_\mathbf{b}) & = & \mathbb{S}[r]_{code}
\end{eqnarray}

@@

We must now define the $\Xi$ function. In most practical implementations this will be modelled as an iterative progression of the pair comprising the full system state, $\mathbb{S}$ and the machine state, $\mathbb{M}$. Formally, we define it recursively with a function $X$. This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine. We consider the empty sequence (denoted $[]$ to be non-equal to the empty set, denoted $\varnothing$).

\begin{eqnarray}
\Xi(\mathbb{S}, g, I) & \equiv & X(\mathbb{S}, \mathbb{M}, I) \\
\mathbb{M}_g & \equiv & g \\
\mathbb{M}_{pc} & \equiv & 0 \\
\mathbb{M}_M & \equiv & [0, 0, ...] \\
\mathbb{M}_i & \equiv & 0 \\
\mathbb{M}_S & \equiv & []
\end{eqnarray}
\begin{equation}
X(\mathbb{S}, \mathbb{M}, I) \equiv \begin{cases}
(\mathbb{S}, \mathbb{M}, I, []) & \text{if} \quad Z(\mathbb{S}, \mathbb{M}, I)\\
(O(\mathbb{S}, \mathbb{M}, I), \mathbf{o}) & \text{if} \quad \mathbf{o} \neq \varnothing\\
X(O(\mathbb{S}, \mathbb{M}, I)) & \text{otherwise}\\
\end{cases}
\end{equation}

where
\begin{equation}
\mathbf{o} \equiv H(\mathbb{S}, \mathbb{M}, I)
\end{equation}

The machine state $\mathbb{M}$ is defined as the tuple $(g, pc, M, i, S)$ which are the gas available, the program counter, the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents. The memory contents $\mathbb{M}_M$ are a series of zeroes of size $2^{256}$.

@@

\begin{equation}
w \equiv \begin{cases} I_\mathbf{b}[\mathbb{M}_{pc}] & \text{if} \quad \mathbb{M}_{pc} < \lVert I_\mathbf{b} \rVert \\
\text{\small STOP} & \text{otherwise}
\end{cases}
\end{equation}

@@

The exceptional halting function $Z$ is defined as:
\begin{equation}
Z(\mathbb{S}, \mathbb{M}, I) \equiv \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}) \vee \mathbf{\delta}_w = \varnothing \vee \lVert\mathbb{M}_S\rVert < \mathbf{\delta}_w
\end{equation}

@@

We define the normal halting function $H$:
\begin{equation}
H(\mathbb{S}, \mathbb{M}, I) \equiv \begin{cases}
H_{\text{\tiny RETURN}}(\mathbb{M}) & \text{if} \quad w = \text{\small RETURN} \\
[] & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SUICIDE} \} \\
\varnothing & \text{otherwise}
\end{cases}
\end{equation}

@@

Stack items are added or removed from the left-most, lower-indexed portion of the series; all other items remain unchanged:
\begin{eqnarray}
O(\mathbb{S}, \mathbb{M}, I) & \equiv & (\mathbb{S}', \mathbb{M}', I) \\
\Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
\lVert\mathbb{M}'_{S}\rVert & \equiv & \lVert\mathbb{M}_T\rVert + \Delta \\
\quad \forall x \in [\mathbf{\alpha}_w, \lVert\mathbb{M}'_S\rVert): \mathbb{M}'_S[x] & \equiv & \mathbb{M}_S[x+\Delta]
\end{eqnarray}

@@

In general, we assume the memory and system state don't change:
\begin{eqnarray}
\mathbb{M}'_M & \equiv & \mathbb{M}_M \\
\mathbb{M}'_i & \equiv & \mathbb{M}_i \\
\mathbb{S}' & \equiv & \mathbb{S}
\end{eqnarray}

@@

As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined:
\begin{eqnarray}
\Pi(\mathbb{S}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\mathbb{S}, T_0), T_1) ...) \\
B & \equiv & (H, [ T_0, T_1, ... ], \mathbf{U})
\end{eqnarray}

@@

\begin{tabular*}{\columnwidth}[h]{lrl}
\toprule
Name & Value & Description* \\
\midrule
$G_{step}$ & 1 & Default amount of gas to pay for execution cycle. \\
$G_{stop}$ & 0 & Nothing paid for the {\small STOP} operation. \\
$G_{sha3}$ & 20 & Paid for a {\small SHA3} operation. \\
$G_{sload}$ & 20 & Paid for a {\small SLOAD} operation. \\
$G_{sstore}$ & 100 & Paid for a normal {\small SSTORE} operation (doubled of waived sometimes). \\
$G_{balance}$ & 20 & Paid for a {\small BALANCE} operation. \\
$G_{create}$ & 100 & Paid for a {\small CREATE} operation and for a contract creation transaction. \\
$G_{call}$ & 20 & Paid for a {\small CALL} operation and for a message call transaction. \\
$G_{memory}$ & 1 & Paid for every additional word when expanding memory. \\
$G_{txdata}$ & 5 & Paid for every byte of data or code for a transaction. \\
\bottomrule
\end{tabular*}

@@

As previously specified in section \ref{ch:model}, these definitions take place in the final context there. In particular we assume $O$ is the EVM state-progression function and define the terms pertaining to the next cycle's state $\mathbb{S}'$ $\mathbb{M}'$ such that:

\begin{equation}
O(\mathbb{S}, \mathbb{M}, I) \equiv (\mathbb{S}', \mathbb{M}', I)
\end{equation}

When interpreting 256-bit binary values as integers the representation is big-endian.

When a 32-byte machine datum is converted to and from as a 160-bit address or hash, the leftwards (high-order for BE) 20 bytes are used and the right most 12 are discarded or filled with zeroes.

The general gas cost function is defined as:

\begin{equation}
C(\mathbb{S}, \mathbb{M}) \equiv G_{memory}(\mathbb{M}'_i - \mathbb{M}_i) + \begin{cases}
C_\text{\tiny SSTORE}(\mathbb{S}, \mathbb{M}) & \text{if} \quad o = \text{\small SSTORE} \\
G_{call} & \text{if} \quad o = \text{\small CALL}\\
G_{create} & \text{if} \quad o = \text{\small CREATE}\\
G_{sha3} & \text{if} \quad o = \text{\small SHA3}\\
G_{sload} & \text{if} \quad o = \text{\small SLOAD}\\
G_{balance} & \text{if} \quad o = \text{\small BALANCE}\\
G_{stop} & \text{if} \quad o = \text{\small STOP}\\
G_{step} & \text{otherwise}
\end{cases}
\end{equation}

@@

Here given are the various exceptions to the state transition rules given in section \ref{ch:model} specified for each instruction, together with the additional instruction-specific definitions of $J$ and $C$. For each instruction, also specify is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from stack.

>>>>

\begin{eqnarray}
\mathbb{S}_{t+1} & \equiv & \Pi(\mathbb{S}_t, B) \\
B & \equiv & (..., (T_0, T_1, ...) ) \\
\Pi(\mathbb{S}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\mathbb{S}, T_0), T_1) ...)
\end{eqnarray}

@@

\section{Conventions}\label{ch:conventions}

I use a number of typographical conventions for the formal notation, some of which are quite particular to the present work:

Highly structured, `top-level', state values, are denoted with black-board type. They fall into two main sets, those of world-state, which are denoted $\mathbb{S}$ (or a variant thereupon) and those of machine-state, $\mathbb{M}$.

Functions operating on highly structured values are denoted with an upper-case greek letter, \eg $\Upsilon$, the Ethereum state transition function.

For most functions, an uppercase letter is used, e.g. $C$, the general cost function. These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation. For specialised and possibly externally defined functions, I may format as typewriter text, \eg the SHA3 hash function is denoted $\texttt{SHA3}$.

Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction. This symbol may, if accoringly defined, be subscripted to refer to an individual component, \eg $T_s$, denotes the timestamp of said transaction. The form of the subscript is used to denote its type; \eg bold subscripts refer to sequences

Scalars and fixed-size byte vectors are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce. Those with a partcularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.

Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte-sequence given as the output data of a message call. For particularly important values, an bold uppercase letter may be used.

Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\mathbb{M}_\mathbf{s}[0]$ denotes the first item on the machine's stack. For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\mathbb{M}_\mathbf{m}[0..31]$ denotes the first 32-byte word of the machine's memory.

In the case of the global state $\mathbb{S}$, which is a sequence of accounts, themselves tuples, the square brackets are used to reference an individual account.

When considering variants of existing values, I follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c. On very particular occasions, in order to maximise readability and only if unambiguous in meaning, I may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.

@@

\begin{description}
\item[nonce] A scalar value equal to the number of transactions sent from this address, or, in the case of contract accounts, the number of contract-creations made by this account. For account of address $a$ in state $\mathbb{S}$, this would be formally denoted $\mathbb{S}[a]_n$.
\item[balance] A scalar value equal to the number of wei owned by this address. Formally denoted $\mathbb{S}[a]_b$.
\item[stateRoot] A 256-bit hash equal to the root node of a further trie structure that encodes the storage contents of the contract. Though a separate data structure, this trie is still stored using the same underlying state database. This trie takes the form as a simple mapping between domains of 256-bit values. The storage, modelled as a $2^{256}$ size sequence of 256-bit values is formally denoted $\mathbb{S}[a]_\mathbf{s}$.
\item[codeHash] The hash of the EVM code of this contract---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction. All such code fragments are contained in the state database under their corresponding hashes for later retrieval. This hash is formally denoted $\mathbb{S}[a]_c$, and thus the code is denoted as $c$, given that $c = \texttt{\small SHA3}(\mathbb{S}[a]_c)$.
\end{description}

@@

the validity is determined as:
\begin{eqnarray}
S(T) & \neq & \varnothing \\ \wedge \quad \mathbb{S}[S(T)] & \neq & \varnothing \\ \wedge \quad T_n & = & \mathbb{S}[S(T)]_n \\ \wedge \quad g_0 & \leq & T_g \\ \wedge \quad v_0 & \leq & \mathbb{S}[S(T)]_b
\end{eqnarray}

@@

We define the rollback state $\mathbb{S}_0$:
\begin{eqnarray}
\mathbb{S}_0 & \equiv & \mathbb{S} \quad \text{except:} \\
\mathbb{S}_0[S(T)]_b & \equiv & \mathbb{S}[S(T)]_b - v_0 \\
\mathbb{S}_0[S(T)]_n & \equiv & \mathbb{S}[S(T)]_n + 1
\end{eqnarray}

@@

After the message call or contract creation is processed, the state is finalised by refunding $g'$, the remaining gas, to the sender at the original rate. The Ether for the gas that was actually used is given to the miner, whose address is specified as the coinbase of the present block $B$. So we define the final state $\mathbb{S}'$ in terms of the provisional state $\mathbb{S}_P$:
\begin{eqnarray}
\mathbb{S}' & \equiv & \mathbb{S}_P \quad \text{except} \\
\mathbb{S}'[s]_b & \equiv & \mathbb{S}_P[s]_b + g' T_p \\
\mathbb{S}'[m]_b & \equiv & \mathbb{S}_P[m]_b + (T_g - g' - g_d) T_p \\
m & \equiv & {B_H}_b
\end{eqnarray}

%In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
%\begin{eqnarray}
%\mathbb{S}'[s]_b & \equiv & \mathbb{S}_P[s]_b + g
%\end{eqnarray}

@@

The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the SHA3 256-bit hash of the code, thus the mutated state becomes $\mathbb{S}^*$:
\begin{eqnarray}
\mathbb{S}^* & \equiv & \mathbb{S} \quad \text{except:} \\
\mathbb{S}^*[A(s, \mathbb{S}[s]_n)] & \equiv & ( 0, v, \varnothing, \mathtt{\tiny SHA3}(\mathbf{b}) )
\end{eqnarray}

@@

We define $\mathbb{S}_1$, the rollback state as the original state but with the value transferred to the recipient:
\begin{eqnarray}
\mathbb{S}_1 & \equiv & \mathbb{S} \quad \text{except:} \\
\mathbb{S}_1[r]_b & \equiv & \mathbb{S}_1[r]_b + v
\end{eqnarray}

If the recipient account contains no code to be executed (i.e. it's not a contract), then we define the final state as being equivalent to the rollback state. In the case that the recipient contains code to be executed (i.e. that the account is a contract), then the contract's body code (identified as the fragment whose SHA3 hash is $\mathbb{S}[r]_c$) is executed according to the execution model (see section \ref{ch:model}). Just as with contract creation, if the execution halts due to an exhausted gas supply, then no gas is refunded to the caller and the state is reverted to the point immediately prior to code execution.
\begin{eqnarray}
\mathbb{S}_P & \equiv & \begin{cases}
\mathbb{S}_1 & \text{if} \quad \mathbb{S}_1[r]_c = \varnothing \\
\mathbb{S}_1 & \text{if} \quad \mathbb{S}^{**} = \varnothing \\
\mathbb{S}^{**} & \text{otherwise}
\end{cases} \\
(\mathbb{S}^{**}, g', \mathbf{o}) & \equiv & \Xi(\mathbb{S}_1, g, I) \\
I_a & \equiv & a \\
I_o & \equiv & o \\
I_p & \equiv & p \\
I_\mathbf{d} & \equiv & \mathbf{d} \\
I_s & \equiv & s \\
I_v & \equiv & v \\
\text{Let} \quad \mathtt{\tiny SHA3}(I_\mathbf{b}) & = & \mathbb{S}[r]_c
\end{eqnarray}

@@

We must now define the $\Xi$ function. In most practical implementations this will be modelled as an iterative progression of the pair comprising the full system state, $\mathbb{S}$ and the machine state, $\mathbb{M}$. Formally, we define it recursively with a function $X$. This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine. The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$.

\begin{eqnarray}
\Xi(\mathbb{S}, g, I) & \equiv & X(\mathbb{S}, \mathbb{M}, I) \\
\mathbb{M}_g & \equiv & g \\
\mathbb{M}_{pc} & \equiv & 0 \\
\mathbb{M}_\mathbf{m} & \equiv & (0, 0, ...) \\
\mathbb{M}_i & \equiv & 0 \\
\mathbb{M}_\mathbf{s} & \equiv & ()
\end{eqnarray}
\begin{equation}
X(\mathbb{S}, \mathbb{M}, I) \equiv \begin{cases}
(\mathbb{S}, \mathbb{M}, I, ()) & \text{if} \quad Z(\mathbb{S}, \mathbb{M}, I)\\
O(\mathbb{S}, \mathbb{M}, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
X(O(\mathbb{S}, \mathbb{M}, I)) & \text{otherwise}\\
\end{cases}
\end{equation}

\begin{eqnarray}
\mathbf{o} & \equiv & H(\mathbb{M}, I) \\
(a, b, c) \cdot d & \equiv & (a, b, c, d)
\end{eqnarray}

The machine state $\mathbb{M}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter, the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents. The memory contents $\mathbb{M}_\mathbf{m}$ are a series of zeroes of size $2^{256}$.

@@

\begin{equation}\label{eq:currentoperation}
w \equiv \begin{cases} I_\mathbf{b}[\mathbb{M}_{pc}] & \text{if} \quad \mathbb{M}_{pc} < \lVert I_\mathbf{b} \rVert \\
\text{\small STOP} & \text{otherwise}
\end{cases}
\end{equation}

@@

The exceptional halting function $Z$ is defined as:
\begin{equation}
Z(\mathbb{S}, \mathbb{M}, I) \equiv \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}, I) \vee \mathbf{\delta}_w = \varnothing \vee \lVert\mathbb{M}_\mathbf{s}\rVert < \mathbf{\delta}_w
\end{equation}

@@

We define the normal halting function $H$:
\begin{equation}
H(\mathbb{M}, I) \equiv \begin{cases}
H_{\text{\tiny RETURN}}(\mathbb{M}) & \text{if} \quad w = \text{\small RETURN} \\
() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SUICIDE} \} \\
\varnothing & \text{otherwise}
\end{cases}
\end{equation}

@@

Stack items are added or removed from the left-most, lower-indexed portion of the series; all other items remain unchanged:
\begin{eqnarray}
O(\mathbb{S}, \mathbb{M}, I) & \equiv & (\mathbb{S}', \mathbb{M}', I) \\
\Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
\lVert\mathbb{M}'_\mathbf{s}\rVert & \equiv & \lVert\mathbb{M}_\mathbf{s}\rVert + \Delta \\
\quad \forall x \in [\mathbf{\alpha}_w, \lVert\mathbb{M}'_\mathbf{s}\rVert): \mathbb{M}'_\mathbf{s}[x] & \equiv & \mathbb{M}_\mathbf{s}[x+\Delta]
\end{eqnarray}

@@

In general, we assume the memory and system state don't change:
\begin{eqnarray}
\mathbb{M}'_\mathbf{m} & \equiv & \mathbb{M}_\mathbf{m} \\
\mathbb{M}'_i & \equiv & \mathbb{M}_i \\
\mathbb{S}' & \equiv & \mathbb{S}
\end{eqnarray}

@@

As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined:
\begin{eqnarray}
\Pi(\mathbb{S}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\mathbb{S}, T_0), T_1) ...) \\
B & \equiv & (H, (T_0, T_1, ...), \mathbf{U})
\end{eqnarray}

@@

The fee schedule $G$ is a tuple of 10 scalar values corresponding to the relative costs, in gas, of a number of abstract operations that a transaction may effect.

\begin{tabular*}{\columnwidth}[h]{lrl}
\toprule
Name & Value & Description* \\
\midrule
$G_{step}$ & 1 & Default amount of gas to pay for execution cycle. \\
$G_{stop}$ & 0 & Nothing paid for the {\small STOP} operation. \\
$G_{sha3}$ & 20 & Paid for a {\small SHA3} operation. \\
$G_{sload}$ & 20 & Paid for a {\small SLOAD} operation. \\
$G_{sstore}$ & 100 & Paid for a normal {\small SSTORE} operation (doubled or waived sometimes). \\
$G_{balance}$ & 20 & Paid for a {\small BALANCE} operation. \\
$G_{create}$ & 100 & Paid for a {\small CREATE} operation and for a contract creation transaction. \\
$G_{call}$ & 20 & Paid for a {\small CALL} operation and for a message call transaction. \\
$G_{memory}$ & 1 & Paid for every additional word when expanding memory. \\
$G_{txdata}$ & 5 & Burnt (discarded from the system) for every byte of data or code for a transaction. \\
\bottomrule
\end{tabular*}

@@

When interpreting 256-bit binary values as integers, the representation is big-endian.

When a 256-bit machine datum is converted to and from a 160-bit address or hash, the rightwards (low-order for BE) 20 bytes are used and the left most 12 are discarded or filled with zeroes, thus the integer values (when the bytes are interpreted as big-endian) are equivalent.

\subsection{Gas Cost}

The general gas cost function, $C$, is defined as:

\begin{eqnarray}
C(\mathbb{S}, \mathbb{M}, I) & \equiv & G_{memory}(\mathbb{M}'_i - \mathbb{M}_i) + \begin{cases}
C_\text{\tiny SSTORE}(\mathbb{S}, \mathbb{M}) & \text{if} \quad w = \text{\small SSTORE} \\
G_{call} & \text{if} \quad w = \text{\small CALL}\\
G_{create} & \text{if} \quad w = \text{\small CREATE}\\
G_{sha3} & \text{if} \quad w = \text{\small SHA3}\\
G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
G_{stop} & \text{if} \quad w = \text{\small STOP}\\
G_{step} & \text{otherwise}
\end{cases} \\
w & \equiv & \begin{cases} I_\mathbf{b}[\mathbb{M}_{pc}] & \text{if} \quad \mathbb{M}_{pc} < \lVert I_\mathbf{b} \rVert \\
\text{\small STOP} & \text{otherwise}
\end{cases}
\end{eqnarray}

@@

As previously specified in section \ref{ch:model}, these definitions take place in the final context there. In particular we assume $O$ is the EVM state-progression function and define the terms pertaining to the next cycle's state $(\mathbb{S}', \mathbb{M}')$ such that:

\begin{equation}
O(\mathbb{S}, \mathbb{M}, I) \equiv (\mathbb{S}', \mathbb{M}', I) \quad \text{with exceptions, as noted}
\end{equation}

Here given are the various exceptions to the state transition rules given in section \ref{ch:model} specified for each instruction, together with the additional instruction-specific definitions of $J$ and $C$. For each instruction, also specified is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from stack, as defined in section \ref{ch:model}.

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{0s: Stop and Arithmetic Operations}} \\
\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$.} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
&&&& $\mathbb{M}'_R = []$ \\
\midrule
0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_\mathbf{s}[0] + \mathbb{M}_\mathbf{s}[1]$ \\
\midrule
0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_\mathbf{s}[0] \times \mathbb{M}_\mathbf{s}[1]$ \\
\midrule
0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_\mathbf{s}[0] - \mathbb{M}_\mathbf{s}[1]$ \\
\midrule
0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_\mathbf{s}[1] = 0\\ \lfloor\mathbb{M}_\mathbf{s}[0] \div \mathbb{M}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
\midrule
0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation. \\
&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_\mathbf{s}[1] = 0\\ \lfloor\mathbb{M}_\mathbf{s}[0] \div \mathbb{M}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
\midrule
0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_\mathbf{s}[1] = 0\\ \mathbb{M}_\mathbf{s}[0] \bmod \mathbb{M}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
\midrule
0x07 & {\small SMOD} & 2 & 1 & Modulo remainder operation. \\
&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_\mathbf{s}[1] = 0\\ \mathbb{M}_\mathbf{s}[0] \bmod \mathbb{M}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
\midrule
0x08 & {\small EXP} & 2 & 1 & Exponential operation. \\
&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_\mathbf{s}[0] ^ {\mathbb{M}_\mathbf{s}[1] }$ \\
\midrule
0x09 & {\small NEG} & 1 & 1 & Negation operation. \\
&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv -\mathbb{M}_\mathbf{s}[0]$ \\
&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
\midrule
0x0a & {\small LT} & 2 & 1 & Less-than comparision. \\
&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_\mathbf{s}[0] < \mathbb{M}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x0b & {\small GT} & 2 & 1 & Greater-than comparision. \\
&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_\mathbf{s}[0] > \mathbb{M}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x0c & {\small EQ} & 2 & 1 & Equality comparision. \\
&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_\mathbf{s}[0] = \mathbb{M}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x0d & {\small NOT} & 1 & 1 & Simple not operator. \\
&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{10s: Bitwise Logic Operations}} \\
\multicolumn{5}{l}{$\mathbb{M}_\mathbf{s}[0]_i$ gives the $i$th bit (counting from zero) of $\mathbb{M}_\mathbf{s}[0]$} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x10 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
&&&& $\forall i \in [0..255]: \mathbb{M}'_\mathbf{s}[0]_i \equiv \mathbb{M}_\mathbf{s}[0]_i \wedge \mathbb{M}_\mathbf{s}[1]_i$ \\
\midrule
0x11 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
&&&& $\forall i \in [0..255]: \mathbb{M}'_\mathbf{s}[0]_i \equiv \mathbb{M}_\mathbf{s}[0]_i \vee \mathbb{M}_\mathbf{s}[1]_i$ \\
\midrule
0x12 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
&&&& $\forall i \in [0..255]: \mathbb{M}'_\mathbf{s}[0]_i \equiv \mathbb{M}_\mathbf{s}[0]_i \oplus \mathbb{M}_\mathbf{s}[1]_i$ \\
\midrule
0x13 & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
&&&& $\forall i \in [0..7]: \mathbb{M}'_\mathbf{s}[0]_i \equiv \begin{cases} \mathbb{M}_\mathbf{s}[1]_{(i + 8\mathbb{M}_\mathbf{s}[0])} & \text{if} \quad \mathbb{M}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
&&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
\bottomrule
\end{tabular*}
